---
title: "p8105_hw3_hbm2126"
output: github_document
date: "2022-10-13"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r reading data}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)

```

### Describing the dataset

The "instacart" dataset contains information on various variables that pertain to orders placed on instacart. There are a couple of identifier variables for the order, product, and the user that placed the order. Other variables include the product name, the aisle, the department the product is from, and days since prior order. There are `r nrow(instacart)` observations and `r ncol(instacart)` columns. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

### Answering questions using the dataset

There are 134 aisles. Fresh fruits and fresh vegetables are the two aisles that most items are ordered from. 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Below is a plot that shows the number of items ordered in each aisle, arranged in an ascending order.

```{r plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The table below shows the 3 most popular items in the following aisles: packaged vegetables & fruits, baking ingredients, and dog food and care, and how many times those items were ordered. 

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()

```

The table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.


```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```






