---
title: "Data Science Homework 3"
output: github_document
date: "2022-10-17"
---

Hemangi Modi (UNI: hbm2126)

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r reading data}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)

```

### Describing the dataset

The "instacart" dataset contains information on various variables that pertain to orders placed on instacart. There are a couple of identifier variables for the order, product, and the user that placed the order. Other variables include the product name, the aisle, the department the product is from, and days since prior order. There are `r nrow(instacart)` observations and `r ncol(instacart)` columns. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

### Answering questions using the dataset

There are 134 aisles. Fresh fruits and fresh vegetables are the two aisles that most items are ordered from. 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Below is a plot that shows the number of items ordered in each aisle, arranged in an ascending order.

```{r plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The table below shows the 3 most popular items in the following aisles: packaged vegetables & fruits, baking ingredients, and dog food and care, and how many times those items were ordered. 

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>% 
  knitr::kable()

```

The table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.


```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

# Problem 2

```{r importing and cleaning data}
accel_df = 
  read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_measure") %>% 
  select(week, day_id, everything()) %>% 
  mutate(
    weekend_weekday = case_when(day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "Weekday", day %in% c("Saturday","Sunday") ~ "Weekend"), 
    minute = as.integer(minute)
    ) 
```

### Describing resulting dataset

After cleaning the dataset in the code chunk above, the resulting dataset has `r nrow(accel_df)` observations and `r ncol(accel_df)` variables. The dataset contains the following variables: what day and week the observation is from, the minute-interval associated with the observation, the actual physical activity measure, and a variable that indicates whether this observation was recorded on a weekday or a weekend. 



The table below shows total activity per day across each one of the 5 weeks. 

```{r}
total_activity = 
  accel_df %>% 
  group_by(day, week) %>% 
  summarize(sum(activity_measure)) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "sum(activity_measure)"
  ) %>% 
  select(week, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable(digits = 2)
```

The plot below shows the 24-hour activity time courses for each day. Some patterns that are apparent from this plot: Wednesday has some of the highest activity measures in the week. Most activity starts around the 400th minute, on all days. 

```{r}
accel_df %>% 
  ggplot(aes(x = minute, y = activity_measure, color = day)) +
  geom_line()
```

# Problem 3

```{r reading in data}
data("ny_noaa")

ny_noaa = 
  ny_noaa %>% 
  as_tibble(ny_noaa)

```

### Describing the dataset

The NY NOAA dataset has `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables. This dataset contains weather data for all New York state weather stations from ` r min(ny_noaa$date)` to `r max(ny_noaa$date)` The variables in this dataset include the station ID, the date of observation, precipitation (in tenths of mm), snowfall (in mm), snow depth (in mm), and maximum and minimum temperatures (in tenths of degrees). This dataset contains a lot of missing values: more precisely, there are `r sum(is.na(ny_noaa))` missing values in this data.

```{r cleaning data}
ny_noaa_tidy = 
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>%
  mutate(
    tmax = as.integer(tmax),
    tmin = as.integer(tmin),
    tmax = tmax/10,
    tmin = tmin/10,
    prcp = prcp/10
    ) 
```

### Answering questions about the dataset (after tidying)

The most commonly observed value for snowfall is `r names(which.max(table(ny_noaa$snow)))` mm. I think this makes sense because it only snows during the winter months and even in the winter months, it does not snow everyday. Therefore, for most days in the year, we would expect no snowfall.


### Making plots

The plot below shows the average max temperature in January and July in each station across years. The average max temperature is much higher in July than in January, which makes sense. The average max temperature in July has been gradually increasing over the years. 

```{r}
avg_tmax_plot = 
  ny_noaa_tidy %>% 
  group_by(id, year, month) %>% 
  filter(month %in% c("01", "07")) %>% 
  mutate(
    month = recode(month, "01" = "Jan", "07" = "July")
  ) %>% 
  summarize(avg_tmax = mean(tmax)) %>% 
  ggplot(aes(x = year, y = avg_tmax, group = id, color = month)) +
  geom_line() +
  facet_grid(~month) +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
avg_tmax_plot
  
```

```{r}
tmax_tmin_plot = 
  ny_noaa_tidy %>% 
  ggplot(aes(x = tmax, y = tmin, color = id)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")
```

```{r}
snowfall_dist = 
  ny_noaa_tidy %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = snow, fill = id)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")
```

```{r}
snowfall_dist = 
  ny_noaa_tidy %>% 
  filter(snow > 0 & snow < 100) %>%
  ggplot(aes(x = snow, y = year)) + 
  geom_density_ridges(scale = .85)
```



```{r}
combined_plot = tmax_tmin_plot + snowfall_dist
combined_plot
```

